An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.

### 2021-04-20 manuscript/asa_template/asa_template.Rmd
update to asa template

---
title: "On the use of 'boilerplate' statistical analysis sections?"
author: "Nicole White, Thiru Balasubramaniam, Richi Nayak, Adrian Barnett"
date: "21/05/2020"
output: bookdown::pdf_document2
toc: false
link-citations: yes
bibliography: references.bib
subtitle: An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyr)
library(pander)
library(ggplot2)
library(ggpubr)
library(knitr)
library(flextable)
library(tidytext)
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('table.emphasize.rownames', FALSE)
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('table.split.table', Inf)
panderOptions('table.split.cells', Inf)
panderOptions('big.mark', ',')
panderOptions('keep.line.breaks', TRUE)
panderOptions('table.alignment.default','left')

g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal',
                           axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))
```

@@

There is guideance for researchers on how to write up their statistical methods and results.
The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019]. 
More detailed guideance is given by the SAMPL and EQUATOR guidelines [@Lang2013;@Altman2016] with the latter covering all apsects of the paper. Both of these guidelines were led by Doug Altman, who spoke often and for many years about the need for better statistical reporting. 
The awareness and use of these guidelines could be improved. There were 256 Google Scholar citations to the SAMPL paper (as at 15 March 2021) which is a good citation statistic for most papers, but is low considering the millions of papers that use statistical analysis.
<!--- Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked. --->

@@

## Data sources

We used two openly available data sources to find statistical methods sections.

### Public Library of Science (PLOS ONE)

@@

Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results. 
<!--- Reporting guidelines are available from the EQUATOR network which has developed guidelines for study designs commonly used in health research, and these guidelines were developed to tackle poor statistical application and reporting in health and medical journals [@Altman2016]. --->

@@

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in the article. Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume and subject classification(s).
<!--- and total article views since publication.--->

_Step 2_: Partial matching on section headings. Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s). Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’. To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]

### Australia and New Zealand Clinical Trials Registry (ANZCTR)

The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered. All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au). 
<!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
Details required for registration follow a standardised template [@ANZCTR], which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes. The information provided must be in English. Studies are not peer reviewed. 
<!---Upon successful registration, the trial is assigned a unique registration number for use in future reporting in scientific publications. --->

@@

All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. 
We used all the text available in the "Statistical methods" section. We also collated basic information about the study including the study type (interventional or observational), submission date, number of funders and target sample size. These variables were chosen as we believed they might influence the completeness of the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.
Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
Some studies were first registered on the alternative trial database _clinicaltrials.gov_ and then also posted to ANZCTR. We excluded these studies because they almost all had no completed statistical methods section as this section is not included in _clinicaltrials.gov_.

## Full-text processing

@@

## Clustering algorithm

@@

## Missing statistical methods sections

@@

```{r}
load('total.records.plos.rda')
n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
perc.include = round(100*n.include/n.api)
min.perc.stats = total.records[[2]] %>% filter(percentage==min(percentage)) %>% select(volume,percentage)
max.perc.stats = total.records[[2]] %>% filter(percentage==max(percentage)) %>% select(volume,percentage)
```

```{r}
load('wordcount.plos.rda')
n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
p.extremes = wordcounts %>% summarise(morethan500 = round(100*mean(words>500)),lessthan50=round(100*mean(words<50)))
n.extremes = wordcounts %>% summarise(morethan500 = sum(words>500),lessthan50=sum(words<50))
```

```{r,fig.width=10,fig.height=12,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}
to_plot = total.records[[2]] %>% select(-percentage) %>% gather(variable,value,-volume) %>%
  mutate(variable=factor(variable,levels=c('Retrieved from search','Included in analysis')))

g1 = ggplot(to_plot,aes(x=volume,y=value,fill=variable))+geom_bar(stat='identity',position='dodge',width=0.9,colour='black')+scale_y_continuous('Total records',breaks=seq(0,20000,2500))+
  scale_x_continuous('Volume',breaks=1:15)+scale_fill_grey(guide = guide_legend(reverse = TRUE) )+g.theme+theme(legend.title = element_blank())

g2 = ggplot(wordcounts,aes(y=words,x=volume,group=volume))+geom_boxplot()+
  scale_x_continuous('Volume',breaks=1:14)+scale_y_log10('Word count (log-10 transformed)',breaks=c(0,10,100,200))+g.theme

load('stats_section_subject.classifications.rda')
stats_section_subject = mutate(stats_section_subject,
top3 = ifelse(subject_level_number %in% 1:3,1,0))
to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')

to_plot.bar = to_plot %>% arrange(-n) %>% select(-n) %>% gather('Top 3 subject classification',value,-subject_level_name) %>%
mutate(subject_level_name=factor(subject_level_name,levels=rev(unique(subject_level_name))),
'Top 3 subject classification'=factor(`Top 3 subject classification`,levels = c('No','Yes')))

g3 = ggplot(to_plot.bar,aes(x=subject_level_name,y=value/10000,fill=`Top 3 subject classification`))+geom_bar(stat='identity',colour='black',width=0.75)+scale_y_continuous('Total records (x 10,000)',breaks=seq(0,12,1))+xlab('')+coord_flip()+scale_fill_manual(values = c('grey75','grey15'))+ theme_bw() + theme(legend.position = c(0.8,0.2),
legend.direction = 'horizontal',axis.text = element_text(size=12),axis.text.x = element_text(size=12),axis.text.y = element_text(size=12))

ggarrange(g1,g2,g3,nrow=3,labels=LETTERS[1:3])

@@

Search results varied by journal volume (Figure \@ref(fig:plos-n)A). The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`), corresponding to years 2013 and 2014. This trend aligned with the total number of papers published in _PLOS ONE_ over the same period. The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).

The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \@ref(fig:plos-n)B). `r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more. `r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.

All papers included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications
(Figure \@ref(fig:plos-n)C).

```{r,fig.width=10,fig.cap="\\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters"}

include_graphics("figures/plosone.10topics.jpg")

@@

```{r}
load('plos.results.10topics.rda')

@@

___
**Box 1: Examples of boilerplate text from _PLOS ONE_ (Statistical software)**

Topic 3

* *graphpad prism (graphpad software, san diego, ca) was used for all analyses.*
* *all statistical analysis was performed using the graphpad prism software.*
* *statistical analysis were done using graphpad prism software (graphpad software, inc., la jolla, ca).*

Topic 5

* *all statistical analyses were conducted in spss 19 (spss inc., chicago, il, usa).*
* *all statistical analyses were performed using the spss version 15 for windows (spss inc., chicago, il, usa). a p-value less than 0.05 was considered a significant difference.*
* *statistical analysis was performed using the spss 17.0 software package (spss, chicago, il, usa). all data were analyzed using t tests, values of p < 0.05 were considered significant.*

___

Definitions of statistical significance featured strongly in Topic~1 (n = 3,775) and Topic 9 (n = 6,104) combined with hypothesis testing for comparing differences between two groups. Topic~1 reflected applications of Student's t-test assuming a 5% level of statistical significance. Topic~9 referenced similar methods combined with multiple thresholds for declaring statistical significance by asterisk: "\*p<0.05, \*\*p<0.01 and \*\*\*p<0.001", a practice that has been criticised [@Wasserstein2019]. Examples of boilerplate text are in Box~2.

___
**Box 2: Examples of boilerplate text from _PLOS ONE_ (Between-group hypothesis testing, two groups)**

Topic 1

* *student’s t-test was used for statistical analysis. a p value of <0.05 was considered statistically significant.*
* *statistical analysis was performed using student`s t-test. a p-value <0.05 was considered to be statistically significant.*
* *statistical analysis was performed using a two-tailed student’s t-test, and p-values <0.05 were considered statistically significant.*

Topic 9

* *statistical significance was determined by students t-tests. \* p<0.05, \*\* p = <0.01, \*\*\* p<0.001.*
* *statistical analyses were performed using the nonparametric mann-whitney test. differences were considered significant for* \**p<0.05,* \*\* *p<0.01, \*\*\* *p<0.001.*
* *comparisons between two groups were performed by student’s test. statistical significance was defined as \**p<0.05;* \*\* *p<0.01;* \*\*\* *p<0.001.*

___

Group-based hypothesis testing was a recurring theme across topics, with text descriptions varying based on method(s) used (Box 3). Frequently occurring words in Topic 6 (n = 4,746) including 'mean', 'sd' and 'sem' reflected the use of descriptive statistics for summarising continuous variables. Sections associated with this topic appeared to be expanded versions of Topics 1 and 9, showing evidence of boilerplate text in the form of descriptive statistics followed by hypothesis testing using Student's t-test, Mann Whitney U or one-way analysis of variance. One-way analysis of variance also featured strongly in Topic 4 (n = 10,163), with text citing common methods for performing post-hoc multiple comparisons. Top matching sections across these topics did not include definitions of groups being compared.

<!-- Topic 10 (n = 4,664) reported similar methods applied to the analysis of replicates collected under experimental designs. Top matching sections across these three topics did not define the groups being compared. [nw: excluded for now as evidence for boilerplate not as strong]-->

Remaining topics reflected applications of meta-analysis (Topic 7: n = 5,165), regression modelling (Topic 2: n = 19,478) and larger sections that reported data processing and statistical methods within the same section (Topic 8: n = 38,183).

<!-- Top matches for Topic 2 shared features of the 'describe-compare' template seen in other topics, expanded to include details of multivariable regression modelling. Examples of multivariable models were (stepwise) logistic regression and Cox proportional hazards regression. -->

___
**Box 3: Examples of boilerplate text from _PLOS ONE_ (Between-group hypothesis testing, two or more groups)**

Topic 4

* *significant differences were determined using analysis-of-variance anova followed by tukey post-hoc tests for multiple comparisons.*
* *statistical comparisons between study groups were performed using one-way anova test followed by post-hoc multiple comparison with dunnett test. p-value of less-than 0.05 were considered to be statistically significant.*
* *one-way analysis-of-variance anova was used to compare multiple groups followed by bonferroni post-hoc test. statistical significance was set at p less-than 0.05.*

Topic 6

* *all results are expressed as means plus-or-minus standard deviation plus-or-minus sd.*
* *data are presented as mean plus-or-minus sd or mean plus-or-minus sem. comparison of means was performed by anova test.*
* *data are expressed as means plus-or-minus standard error of the mean sem. statistical analysis was performed using the student t-test. p less-than 0.05 was considered significant.*

<!-- Topic 10 -->

<!-- * *all statistical analyses were performed using microsoft excel. significance was assessed by the student t-test using data from at least three independent experiments. all images are representative of at least three independent experiments.* -->

<!-- * *data are presented as the mean plus-or-minus sd of at least three experiments or are representative of experiments repeated at least three times. student t-test two-tailed paired was used to evaluate the statistical significance of results from proliferation assay experiments.* -->

<!-- * *data are presented as the mean plus-or-minus sd. comparisons were made using a one-way anova. all experiments were repeated at least three times and representative experiments are shown.* -->

___

## ANZCTR

We downloaded 28,008 studies. The numbers of excluded studies are shown in Figure~X. 
Of the 12,700 included studies, 9,523 (75%) had a statistical methods section.

The median length of the section was 129 words with an inter-quartile range of 71 to 219 words.
Some methods sections were only one word, including "ANOVA", "t-test", "SPSS" and even "SSPS".

The clustering algorithm found groups that were purely sample size calculations (topic 2), pilot studies (topic 5), safety/tolerability studies (topic 6) and repeated measures ANOVA (topic 10). There were cases where the exact same method section had been re-used in a different study.

We also found evidence of 'boilerplate' sections clustered as topic 3, example text

___
**Box x: Examples of boilerplate text from ANZCTR**

Topic 3

* "Shapiro Wilk test was used as normality test. Continuous variables were compared using Student t-test and Mann-Whitney U test when the data were not normally distributed. Categorical variables were compared using Pearson’s chi-squared test and Fisher’s exact test. Paired data were analyzed using Paired t-test and Wilcoxon signed rank test when data were not normally distributed."
* "Comparisons between categorical variables will be made either using chi square or Fisher exact test. Continuous data will be compared using the Student’s t-test or Mann-Whitney U test. Two sided p values of less than 0.05 will be considered statistically significant."
___

We examined if four study characteristics were associated with a missing statistics section. The odds ratios and 95% credible intervals are in Table~X. Observational studies were less likely to have a missing methods section compared with interventional studies. Missing sections became less likely over time. Studies with more funders and a larger target sample size were less likely to have a missing methods section.

```{r}
table =  read.table(header=TRUE, sep='&', text='
Variable & Odds ratio & 95% CI
Study type = Observational & 0.78 & 0.69, 0.89
Date (per year) & 0.90 & 0.88, 0.91
Number of funders & 0.80 & 0.74, 0.86
Target sample size (per doubling) & 0.90 & 0.88, 0.92')

ftab = flextable(table) %>%
theme_box() %>%
autofit()
ftab
```

* Final sample size

# Discussion

The first line in many statistical analysis sections in _PLOS ONE_ was the software used and some entire sections in ANZCTR only stated the software, implying that the software is the most important detail. As Doug Altman said, "Many people think that all you need to do statistics is a computer and appropriate software" [@Altman1994]. This is far from the truth, and whilst it is important for researchers to mention the software and version used for reproducibility purposes, it is a minor detail compared with detailing what methods were used and why.

A frequent theme in the boilerplate statistical methods is the definition of statistical significance, nearly always using a p-value at the 5% level. This widespread use of statistical significance is troubling giving the bright-line thinking it engenders [@McShane2019] and the common misinterpretations of p-values [@Goodman2008].

Despite the extensive array of statistical tests available, many authors are reporting the same few methods.

One reason these inadequate sections get published is that most journals do not use statistical reviewers, despite empirical evidence showing they improve manuscript quality [@Hardwicke2020].

A related paper has criticised vague statistical methods sections because they deprive readers and reviewers for the opportunity to confirm that the appropriate methods were used [@Weissgerber2018]. These authors checked hundreds of papers using ANOVA and found that 95\% did not contain the information needed to determine what type of ANOVA was performed. This lack of information could well be because the authors used a boilerplate statistical methods section that was missing key details.

If authors shared their code then this would provide an alternative route for checking what statistical methods were used. This is not a perfect solution, as we still want authors to accurately report their methods, but it does increase transparency. However, a recent paper found that code sharing was very low in biomedical papers, with just 2\% of a sample of over 6,000 papers sharing code [@Serghiou2021].

Many researchers are using lazy practice by copying a standard "boilerplate" statistical methods section, likely cut-and-pasting from other researchers or projects. This is a strong sign of the ritualistic practice of statistics where researchers go through the motions rather than using conscientious practice [@Stark2018]. This is concerning because using the wrong statistical methods can reduce the value of study, or worse, invalidate the entire study. These mistakes are avoidable and are wasting of thousands of hours of researchers' time and the time of patients and volunteers. Poor statistical practice is a key driver of the ongoing reproducibility crisis in science [@Ioannidis2014].

## Limitations

We did not check whether papers used the correct methods, and for some simple studies a 'boilerplate' statistical methods might be adequate.

We examined papers where there was a statistics section, and we missed papers that used statistical analysis but did not include a statistical analysis section. Reiterate outcomes of random sample checking here.

We only examined one large journal and one trial registry and hence our results may not be generalisable to all journals or registries, especially those that consistently use a statistical reviewer.

We searched the full text of _PLOS ONE_ papers but not the supporting information which may contain statistical methods sections for some papers. The search terms we used to find statistical methods appeared in the supporting information titles for xxx papers (x%). We did not include the supporting information because it is less structured than the paper and could be in PDF or Word format.

# References

>>>>

---
title: An observational study of papers published in _PLOS ONE_ and studies posted to a trial registry.

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Nicole White
  thanks: The authors gratefully acknowledge ...
  affiliation: School of Public Health and Social Work, QUT

- name: Thiru Balasubramaniam, Richi Nayak
  affiliation: To add

- name: Adrian Barnett
  affiliation: School of Public Health and Social Work, QUT

keywords:
- 3 to 6 keywords
- that do not appear in the title

abstract: |
  The text of your abstract.  200 or fewer words.

bibliography: references.bib
output: rticles::asa_article
fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
library(knitr)
library(flextable)
library(tidytext)

g.theme = theme_bw()+theme(legend.position = 'top',legend.direction = 'horizontal')
```

# Introduction

@@

There is guidance for researchers on how to write up their statistical methods and results.
The International Committee of Medical Journal Editors recommend that researchers should: “Describe statistical methods with enough detail to enable a knowledgeable reader with access to the original data to judge its appropriateness for the study and to verify the reported results” [@ICMJE2019]. 
More detailed guideance is given by the SAMPL and EQUATOR guidelines [@Lang2013;@Altman2016] with the latter covering all apsects of the paper. Both of these guidelines were led by Doug Altman, who spoke often and for many years about the need for better statistical reporting. 
The awareness and use of these guidelines could be improved. There were 256 Google Scholar citations to the SAMPL paper (as at 15 March 2021) which is a good citation statistic for most papers, but is low considering the millions of papers that use statistical analysis.
<!--- Although the general lack of statistical understanding from both authors and reviewers means this recommendation may not be checked. --->

@@

\subsection{Data sources}

We used two openly available data sources to find statistical methods sections, covering peer-reviewed journal articles and registered clinical trial protocols.

\subsubsection{Public Library of Science (PLOS ONE)}
\label{sec:methodsPLOS}

@@

Authors are encouraged to follow published reporting guidelines such as EQUATOR, to ensure that chosen statistical methods are appropriate for the study design, and adequate details are provided to enable independent replication of results.

@@

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in the article. Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume and subject classification(s).

_Step 2_: Partial matching on section headings. Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s). Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings agai

\subsubsection{Australia and New Zealand Clinical Trials Registry (ANZCTR)}
\label{sec:methodsANZCTR}

The ANZCTR was established in 2005 as part of a coordinated global effort to improve research quality and transparency in clinical trials reporting; observational studies can also be registered. All studies registered on ANZCTR are publicly available and can be searched via an online portal (https://www.anzctr.org.au). 
<!--- Researchers responsible for overseeing a clinical trial are expected to register full details before commencing participant enrolment. --->
Details required for registration follow a standardised template [@ANZCTR], which covers participant eligibility, the intervention(s) being evaluated, study design and outcomes. The information provided must be in English. Studies are not peer reviewed.

@@

All studies available on ANZCTR were downloaded on 1 February 2020 in XML format. 
We used all the text available in the "Statistical methods" section. We also collated basic information about the study including the study type (interventional or observational), submission date, number of funders and target sample size. These variables were chosen as we believed they might influence the completeness of the statistical methods section, because we expected larger studies and those with funding to be more complete, and we also were interested in changes over time.

Studies prior to 2013 were excluded as the statistical methods section appeared to be introduced in 2013.
Some studies were first registered on the alternative trial database _clinicaltrials.gov_ and then also posted to ANZCTR. We excluded these studies because they almost all had no completed statistical methods section as this section is not included in _clinicaltrials.gov_.

\subsection{Full-text processing}
\label{sec:methods-cleaning}

@@

nst frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’. To determine the reliability of our chosen filters, we manually reviewed full text data extracted for a random sample of XXX articles that were not matched (File S1).[TODO…finish this thought…]

\subsection{Clustering algorithm}

@@

\subsection{Missing statistical methods sections}

@@

```{r}
load('../total.records.plos.rda')
n.api = total.records[[1]] %>% filter(stage=='API') %>% pull(n) %>% sum()
n.include = total.records[[1]] %>% filter(stage=='Included in analysis') %>% pull(n) %>% sum()
perc.include = round(100*n.include/n.api)
min.perc.stats = total.records[[2]] %>% filter(percentage==min(percentage)) %>% select(volume,percentage)
max.perc.stats = total.records[[2]] %>% filter(percentage==max(percentage)) %>% select(volume,percentage)
```

```{r}
load('../wordcount.plos.rda')
n.words = wordcounts %>% summarise(med = median(words),q1=quantile(words,.25),q3=quantile(words,.75))
p.extremes = wordcounts %>% summarise(morethan500 = round(100*mean(words>500)),lessthan50=round(100*mean(words<50)))
n.extremes = wordcounts %>% summarise(morethan500 = sum(words>500),lessthan50=sum(words<50))
```

![\label{fig:plos-n}A: Search results by _PLOS ONE_ volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis](figures/figure1.jpg)

```{r,fig.height=12,fig.width=6,fig.asp=0.6,fig.cap='\\label{fig:plos-n}Search results by _PLOS ONE_ volume (1st row); word count per statistical methods section included in analysis (n = 111,731; 2nd row); subject classifications assigned to full-text records included in analysis (3rd row)'}

load('../stats_section_subject.classifications.rda')
stats_section_subject = mutate(stats_section_subject,
                               top3 = ifelse(subject_level_number %in% 1:3,1,0))

to_plot = stats_section_subject %>% group_by(subject_level_name) %>% summarise(n=n(),Yes=sum(top3),No=sum(top3==0),.groups='drop')

@@

Search results varied by journal volume (Figure \ref{fig:plos-n}A). The total number of API search results peaked at volumes 8 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`) and 9 (n = `r format(total.records[[2]] %>% filter(volume==8) %>% pull(API),big.mark=',')`), corresponding to years 2013 and 2014. This trend aligned with the total number of papers published in _PLOS ONE_ over the same period. The percentage of records that included a statistical methods section by volume based on our proposed matching criteria varied between `r round(min.perc.stats[['percentage']])`% (volume `r min.perc.stats[['volume']]`) and `r round(max.perc.stats[['percentage']])`% (volume `r max.perc.stats[['volume']]`).

The median length of statistical methods sections was `r n.words[['med']]` words (IQR: `r n.words[['q1']]` to `r n.words[['q3']]` words) (Figure \ref{fig:plos-n}B). `r format(n.extremes[['morethan500']],big.mark=',')` articles (`r p.extremes[['morethan500']]`%) had a statistical methods section of 500 words or more. `r format(n.extremes[['lessthan50']],big.mark=',')` articles (`r p.extremes[['lessthan50']]`%) had a statistical methods section of 50 words or less, equal to the length of this paragraph.

All papers included Biology and life sciences (n = `r format(n.biology,big.mark=',')`), Earth sciences
(n = `r format(n.earth,big.mark=',')`) and/or Computer and information sciences (n = `r format(n.computer,big.mark=',')`) in their top 3 subject classifications
(Figure \ref{fig:plos-n}C).

![\label{fig:plot-10-topics}Topic clouds summarising common terms in _PLOS ONE_ statistical methods sections assuming ten clusters](figures/plosone.10topics.jpg)

@@

```{r}
load('../plos.results.10topics.rda')

@@
